---
output:
  pdf_document: default
  html_document: default
---
# Load the packages

```{r}
library(MASS)
library(mvtnorm)
library(invgamma)
library(ggplot2)
library(here)
library(coda)
library(tidyverse)
library(tidybayes)
library(bayesplot)
source(here::here("R", "convert-to-coda.R"))
```

# Gibbs sampling

## Create the data to train the model

```{r }
set.seed(111)
k <- 5
n <- 100
mu_0 <- rep(5,k)
a <- 1
b <- 1
sigma_sq <- rinvgamma(a,b)
X <- cbind(1,mvrnorm(n, mu_0, diag(sqrt(sigma_sq), nrow = k, ncol = k)))
p = ncol(X)
mean_true_beta <- rep(0, 6)
true_beta <- mvrnorm(1, mean_true_beta, diag(4, nrow = p, ncol = p))
y <- X %*% true_beta + rnorm(n)

plot(y, X %*% true_beta)
```

# Codes

```{r, echo = TRUE, warning=FALSE, message=FALSE}
set.seed(100)
mcmc_Gibbs <- function(niter) {
  
  #Initial values
  
  beta <- matrix(0, niter, p)
  sigma_sq <- rep(0, niter)
  sigma_sq[1] <- 1
  
  # sample beta 
  
  for(i in 2 : niter){
    
    mu <- solve(t(X) %*% (X)) %*% (t(X) %*% (y)) # mean of beta
    Sigma <- solve(t(X) %*% (X)) * sigma_sq[i-1] # covariance matrix for beta
    beta[i,] <- mvrnorm(1, mu, Sigma)
    
    # sample sigma_sq
    
    b_n <- 0.5 * t(y - X %*% beta[i,]) %*% (y - X %*% beta[i,]) + 1
    a_n <- (n/2) + 1
    sigma_sq[i] <- rinvgamma(1, a, rate = b)
    
  }
  return(
    list(
       beta = beta, 
    sigma_sq = sigma_sq
    )
   
  )
    
    
      
}

```

# Output

```{r, echo = FALSE, warning=FALSE, message=FALSE}
N = 5000
out <- mcmc_Gibbs(N)
#plot(out$sigma_sq, ylim = c(-1, 100))



burn_in <- seq(1:2000)
beta_out <- out[[1]][-burn_in, ]
sigma_out <- out[[2]][-burn_in]

```
# Model disgnostics

```{r, echo = FALSE, warning=FALSE}

# Rearranging the mcmc output to comfort with coda package

out_coda <- as.mcmc(convert_to_coda_single_chain(out))
out_coda <- tidy_draws(out_coda) ## convert to tidybay

```
# Trace plots

```{r trace-plot, fig.cap="Trace plot for model parameters", echo = FALSE, warning=FALSE}

out_coda %>%
    mcmc_trace(regex_pars = c("beta", "sigma_sq"),
              facet_args = list(nrow = 3,  ncol = 3))
```

### Other plots illustrating posterior distribution for our model parameters.


```{r,  echo = FALSE, warning = FALSE}
plot_title1 <- ggtitle("Credible intervals for model parameter")
plot_title2 <- ggtitle("Credible intervals for regression coefficints")
title3 <- ggtitle("Posterior density for regression coefficients")

par(mfrow =c(2,2))
out_coda %>%
    mcmc_intervals(regex_pars = c("beta", "sigma_sq")) + plot_title1
out_coda %>%
mcmc_areas(regex_pars = c("beta"),
  prob = 0.8, # 80% intervals
  prob_outer = 0.99, # 99%
  point_est = "mean") + plot_title2

color_scheme_set("blue")
mcmc_dens(out_coda, regex_pars = c("beta")) + title3

```

## Regression model diagnositcs

```{r Diagnosis, echo = FALSE, warning=FALSE, fig.cap = "regression model diagnoses plots"} 

beta_hat <- apply(beta_out, 2, mean) # Posterior mean of beta

y_pred <- X %*% beta_hat

residuals <- y - y_pred

# Fitted density vs true density

plot(density(y), col = "red", main = "Fitted density vs observed density")
lines(density(y_pred), col = "blue")
legend("topright", legend = c("true density",  "fitted density"),
       col = c("red", "blue"), lty=1:2, cex=0.8)

# Residual and fitted values plots

par(mfrow = c(2,2))
hist(residuals, main = "residual histogram")
qqnorm(residuals, main = "residual qqplot")
abline(c(0, 1), col ="blue")
plot( y_pred, residuals, main = "resuduals vs fitted values")
plot(y, y_pred, main = "Observed values vs predicted values")
abline(c(0, 1), col ="blue") ## a line of 45 degree angle



# data_out <- data.frame(y = y, y_pred = y_pred)
# ggplot(data_out) +
#   geom_density(aes(y, color ="observed density")) +
#    geom_density(aes(y_pred, color = "predicted density")) +
#    ggtitle("comparison between observed density and predicted density") 


```   


# Metropolis Hastings


mcmc_lm_mh <- function(niter, beta_tune, sigma2_tune){
    
    set.seed(222)
    
    # Prior distribution
    
    prior <- function(beta, sigma2){  
        
        prior_beta <- dmvnorm(beta, rep(0, p), diag(1, nrow = p, ncol = p), log = T) 
        prior_sigma2  <- dunif(sigma2 , min = 0, max = 30, log = T)  
        return(prior_beta + prior_sigma2)
    }
    
    # Likelihood function
    
    likelihood <- function(beta, sigma2){   
        
        mean_y = X %*% true_beta
        log_likelihood = dnorm(y, mean_y, sd = sqrt(sigma2), log = T)
        return(sum(log_likelihood))   
    }
    
    
    # Posterior
    
    posterior <- function(beta, sigma2){
        
        return(prior(beta, sigma2) + likelihood(beta, sigma2))
        
    }
    
   # Initialize parameters
    
    beta_chain        <- matrix(0,  nrow = niter, ncol = p)
    sigma2_chain      <- rep(0, niter)
    beta_chain[1,]    <- mvrnorm(1, rep(0, p), diag(1, nrow = p, ncol = p)) 
    sigma2_chain[1]   <- rinvgamma(1, shape = 1, rate =1)

    
    # Metropolis-Hastings step
    
    for (i in 2 : niter){ 
        
        proposed_beta <- mvrnorm(1, beta_chain[i-1,], diag(beta_tune))
        proposed_simga2 <- rnorm(1, sigma2_chain[i-1], sigma2_tune)
        
        if(proposed_simga2 < 0){
            proposed_simga2 <- sigma2_chain[i-1]

        } else {

            proposed_simga2 <- proposed_simga2
        }
        
        accep_ratio  <- exp(posterior(proposed_beta, proposed_simga2) - posterior(beta_chain[i-1,], sigma2_chain[i-1]))
              alpha  <- min(1, accep_ratio)
              
        # Include the acceptance ratio
        
        #a <- logical(length = niter)
        
        if (runif(1) < alpha) {
            beta_chain[i, ] <- proposed_beta
            sigma2_chain[i] <- proposed_simga2
        } else {
            beta_chain[i, ] <- beta_chain[i-1, ]
            sigma2_chain[i] <- sigma2_chain[i-1]
        }
        
  #       if (runif(1) < alpha) {
  #           beta_chain[i,] <- proposed_beta 
  #           sigma2_chain[i] <- proposed_simga2
  #           a[i] <- TRUE 
  # } else {
  #           beta_chain[i,] <- beta_chain[i-1,] 
  #           sigma2_chain[i] <- sigma2_chain[i-1] 
  #       }
        
    } 
    return(list(
        beta_chain = beta_chain,
        sigma2_chain = sigma2_chain
        #a = a
    ))
}


# mcmc_lm_mh results

beta_tune <- rep(0.01, p)
sigma2_tune <- 0.01
niter <- 5000
out_mh = mcmc_lm_mh(niter, beta_tune, sigma2_tune)

# Trace plots

out_mh$beta_chain
post_sigma2 <- out_mh$sigma2_chain
post_beta <- out_mh$beta_chain

plot(out_mh$sigma2_chain, type = "l" )
abline(h = true_sigma_sq)
matplot(out_mh$beta_chain, type = "l",  xlab = "iterations", col = c("red",  "blue", "green", "purple", "black"))
abline(h = true_beta, col = c("red",  "blue", "green", "purple", "black"))

# Accepted samples

acceptance <- out_mh$a
accepted_samples <- length(acceptance[acceptance[TRUE]])
accepted_samples

accept_ratio <- 100*(accepted_samples/niter)
accept_ratio

# Or

accept_ratio1 <- 100*mean(as.numeric(out_MH$a))
accept_ratio1
